Draft Exploration

### 2020-03-22 draft_exploration.Rmd
cleaned data and made plot of high school signees to MLB

```{r, include=FALSE}
draft <- read_csv("data/draft_register.csv") %>%
  clean_names()

angels <- data.frame(a = c("Anahaim Angels", "Los Angeles Angels of Anaheim", "California Angels"))

replace_angels <- data.frame(from = c("Anahaim Angels",
                                      "Los Angeles Angels of Anaheim",
                                      "California Angels"), 
                             to = c("Los Angeles Angels"))

draft <- FindReplace(data = draft, "teamname", replaceData = replace_angels,
                     from = "from", to = "to", exact = FALSE)

draft <- draft %>%
  mutate(teamname == ifelse(teamname == "Anahaim Angels", "Los Angeles Angels",
                            ifelse(teamname == "California Angels", "Los Angeles Angels",
                                   ifelse(teamname == "Los Angeles Angels of Anaheim", "Los Angeles Angels",
                                          teamname))))
```

@@

```{r, echo=FALSE}
grow_talent <- draft %>%
  filter(source == "H" & signed == "Y") %>%
  mutate(decade = ifelse(year %in% 1970:1979, 1970,
                                 ifelse(year %in% 1980:1989, 1980,
                                        ifelse(year %in% 1990:1999, 1990,
                                               ifelse(year %in% 2000:2009, 2000,
                                                      ifelse(year %in% 2010:2019, 2010, NA)))))) %>%
  filter(year >= 1980 & year <= 2009) %>%
  mutate(mlb_made = ifelse(high_level == "MLB", 1, 0)) %>%
  mutate(ignore = 1) %>%
  filter(teamname != "Los Angeles Angels of Anaheim") %>%
  group_by(teamname) %>%
  summarize(percent_mlb = sum(mlb_made) / sum(ignore))

@@

>>>>

```{r, include=FALSE}
war <- read_csv("data/war.csv")
```

```{r, include=FALSE}

draft <- read_csv("data/draft_register.csv") %>%
  clean_names()

# grouping war data by player and summing career games played and war

war_joined <- war %>%
  group_by(mlb_ID) %>%
  summarize(g = sum(G),
            war = sum(as.numeric(WAR)))

# joined draft and war data by player mlb ID, then mutated teamname to condense
# the Angel's team names into a single team

master <- draft %>%
  left_join(war_joined, by = c("mlbid" = "mlb_ID")) %>%
  mutate(teamname = as.factor(case_when(teamname %in% c("California Angels",
                                                        "Anahaim Angels",
                                                        "Los Angeles Angels of Anaheim") ~
                                                          "Los Angeles Angels",
                                        TRUE ~ teamname)))

# deleting useless columns

master$mlbid_1 <- NULL
master$x27 <- NULL

@@

```{r, echo=FALSE}
grow_talent <- master %>%
  filter(source == "H" & signed == "Y") %>%
  mutate(decade = ifelse(year %in% 1970:1979, 1970,
                                 ifelse(year %in% 1980:1989, 1980,
                                        ifelse(year %in% 1990:1999, 1990,
                                               ifelse(year %in% 2000:2009, 2000,
                                                      ifelse(year %in% 2010:2019, 2010, NA)))))) %>%
  filter(year >= 1980 & year <= 2009) %>%
  mutate(mlb_made = ifelse(high_level == "MLB", 1, 0)) %>%
  mutate(ignore = 1) %>%
  group_by(teamname) %>%
  summarize(percent_mlb = sum(mlb_made) / sum(ignore))

@@

```{r, echo=FALSE}

hs_signed <- master %>%
  filter(source == "H" & signed == "Y") %>%
  mutate(mlb_made = ifelse(high_level == "MLB", 1, 0)) %>%
  mutate(ignore = 1) %>%
  group_by(draft_round) %>%
  filter(year >= 1990 & year <= 2009) %>%
  filter(draft_round <= 40) %>%
  summarize(percent_mlb = sum(mlb_made) / sum(ignore))

ggplot(hs_signed, mapping = aes(draft_round, percent_mlb)) +
  geom_col() +
  theme_economist_white() +
  theme(axis.text.x = element_text(angle = 90)) +
  labs(
    x = "Team",
    y = "Percent of Signed High School Picks to MLB",
    title = "Percent of Signed High School Picks to MLB by Round",
    subtitle = "Data: 1980-2009 Drafts"
  )
